---
title: "Orientation"
author: "Ethan Violette"
date: "4/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE)

library(ggplot2)
library(rstanarm)
library(tidyverse)
```

# Data Import and Formatting

```{r}
orientation.df <- read.csv("n90_pol.csv")

orientation.df$orientation <- as.factor(ifelse(orientation.df$orientation < 3, "conservative", "liberal"))
```

### Preparing the inputs

```{r}
x <- model.matrix(orientation ~ . -1, data = orientation.df)
y <- orientation.df$orientation
```


We'll use the stan_glm function from the rstanarm package to fit a logistic regression model from the Bayesian perspective. To do so, we need specified priors for the predictors **amygdala** and **acc**. A Student t prior with p-1 = 1 degree of freedom and a scale of 2.5 is a good prior when we expect the model coefficients to be close to zero, but accept the possibility of large values. 

**stan_glm** effectively mirrors the generalized linear model function **glm** in R, with extra parameters to specific our aforementioned priors. We'll also set a seed for reproducibility, and add a random effect to account for individual variance.

```{r}
t_prior <- student_t(df = 7, location = 0, scale = 2.5)

post1 <- stan_glm(orientation ~ amygdala + acc, 
                  data = orientation.df, 
                  family = "binomial", 
                  prior = t_prior, 
                  prior_intercept = t_prior,
                  QR = TRUE,
                  seed = 1)
```

```{r}
round(coef(post1), 2)
round(posterior_interval(post1, prob = .9), 2)
```

